Is Partner Modeling Mutual?

### 2014-08-26 project-mutual-modeling.tex
Imported Study 1 from original paper

@@

>>>>

\usepackage[english]{babel}

\usepackage{graphicx}
\graphicspath{{figs/}}

@@

\section{Studies}

\subsection{Study 1: Effect of an awareness tool on \Model{A}{B} in a virtual
game}

\paragraph{Questions}

We studied the impact of an awareness tool on group performance and mutual
modeling. (Nova, Wehrle, Goslin, Bourquin \& Dillenbourg (2006). The availability
of an awareness tool was our independent variable. In previous studies, we
replayed a video of the game to subjects who surprised us by their ability to
remember former states of their mutual model: "I did that because I thought that
you would do that". Hence, this experiment focused the representation of each
other's action plans. During the game, we asked them to anticipate the next
action of their partner as well as to announce their own action. It is clear
asking them a question about their partner increases their attention for the
rest of the game.

\paragraph{Experimental setting}

SpaceMiners is a 3D game that involves two players collecting minerals located
in asteroids. To do so, they shoot drones through the space after choosing their
initial direction and speed. Once launched, the trajectory of drones is only
influenced by the gravity of planets and by trajectory modification tools.
During the experiment, the teams were confronted with three increasingly complex
situations. The experiment was 2 hours long: a 30 minutes tutorial and 3 levels
of 30 minutes. Players uses using a regular joystick and communicated with each
other through an audio channel.

The independent variable was the availability of an awareness tool that shows to
player A  the location and gaze direction of player B. In the awareness
condition, players could switch to the 'scout mode' where they could view what
their partner was looking at. We hypothesize that this would enable subjects to
more accurately infer their teammate's intentions. Each player sat in front of a
distinct computer located in different rooms.

\paragraph{Subjects}

Thirty-six persons participated in this study, all native French speakers. We
constituted 18 pairs of participants (N = 18) who were not familiar with each
other. The pairs were assigned randomly to either the control condition (without
the awareness tool) or the awareness condition (with the awareness tool).

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{image4.png}
        \caption{Snapshot of the SpaceMiner Game}
        \label{spaceminer}
\end{figure}

\paragraph{Variables}

Task performance was measured by the score reached by the two subjects after
three situations. The effort of mutual modeling was measured as the ratio of
time that players would spend in the scout mode (divided by total time), which
is the time during which players are not performing their own actions but
monitoring their partner's actions.

In order to evaluate m°(A,B) during the task, we used two questionnaires (Figure
4) that were displayed during each of the three games, as a transparent layer
appearing over the game display. The first questionnaire concerned the player’s
intended actions. The second questionnaire asked each player about what he
thought her or his partner was intending to do. Some answers were identical in
both questionnaires (like “adjusting a shot”) while others were reversed ("to
guide my partner" versus "to guide me"). We chose this subjective measure of
accuracy (∆ M(A,B), M(B,B)) rather than an objective measure (i.e. model (A,B)
is compared to B's next action) because some the activities proposed by the
questionnaire were not observable by the environment (e.g. establishing a
strategy). We counted M°(A,B,activity,gamei) as the number of common answers
between questionnaires M(A,B) and M(B,B) in each game and computed the average
value a across the 3 games.

Figure 4: M(A,A) and M(A,B) questionnaires in SpaceMiners (translated from French)

\paragraph{Results}

Grounding criterion. The grounding criterion was high: the correlation between
M°(A,B) and task performance was 0.42, p = 0.05. Pairs with an accurate mutual
model reached higher scores. A regression analysis confirmed the positive and
significant relation between group performance and mutual modelling accuracy (β
=54, p = .02).

Study-specific questions. The awareness tool permitted higher group performance,
but it did not improve the accuracy of the mutual model. Since teams were free
to use the awareness tool or not (the 'scout' mode), we performed a post-hoc
split of players depending on how much time they used it. The split point was
the mean of time spent in the scout mode and it led to the constitution of two
groups made up of 12 individuals “short time in scout mode” and 24 individuals
“long time in scout mode”. A two-way analysis of variance conducted on these
contrasted groups revealed that pairs in the awareness condition who spent more
time in the scout mode reached higher levels of M°(A,B)(F = 8.02, p = 0.015). Of
course, a post-hoc split does not support a causal direction. An alternative
explanation could be that good modellers are more social and hence appreciate
the awareness tool.

Symmetry question. We computed intra-class correlation as described by Kenny et
al. (1998) from the answers to the cross-questionnaires. Considering all pairs
in both conditions, we found a positive and significant correlation (r = .38, p
< .05) between M°(A,B) and M°(B,A). Interestingly, this was higher in the
control group (r = 0.44) than in the experimental group (r = 0.24). Actually, ∆
(M°(A,B),M°(B,A)), i.e. the absolute differences between the models accuracy,
was not significantly different with or without the awareness tool ( F [1,13]=
0.144, p > 0.5). This result could be explained by the fact that the players
without awareness tools communicated more.

We cannot answer the triangle questions since this study was conducted with
pairs-

\paragraph{Discussion}

How to interpret a correlation of 0.38 between M°(A,B) and M°(B,A)? If it was
null, modelling would be interpreted as an individual process that varies
according to personality traits or cognitive skills: some persons tend to model
more accurately their peer while others don't care or can simply not. If the
correlation was close to 1, we could conclude that the modelling accuracy is an
emergent property of teams since there would be low probability that, by chance,
low accurate modellers are paired with accurate modellers and conversely. Our
correlation is between these two extremes but the very notion of significant
intra-class correlation tend to indicate that one may not consider peers as
independent subjects, which is an argument for the 'interaction' hypothesis
(H3).
